Learning and comparing functional connectomes between populations

### 2012-12-09 paper.tex
iterate

@@

@@

@@

\begin{figure}
\includegraphics[width=.5\linewidth]{aal_atlas.pdf}%
\hfill%
\includegraphics[width=.5\linewidth]{ncuts_atlas.pdf}%

\includegraphics[width=.5\linewidth]{ho_atlas.pdf}%
\hfill%
\includegraphics[width=.5\linewidth]{sulci_atlas.pdf}%

@@

\caption{
Different full-brain parcellation: the AAL atlas
\cite{tzourio-mazoyer2002a}, regions extracted by Ncuts
\cite{craddock2011}, the Harvard-Oxford atlas, the sucli atlas used in
\cite{varoquaux2010c}, the resting-state networks extracted in
\cite{smith2009} by ICA, and in \cite{varoquaux2011} by sparse dictionary
learning.
\label{fig:parcellations}
}
\end{figure}

@@

% Disclaimer: focus on correlation/second order statistics?

\paragraph{Signal extraction}

Once again separate task from rest (explain the difference between
\emph{ongoing} and \emph{evoked} activity at some point (intro?)

Task: beta time-series, and more recent stuff, \emph{e.g.} by Mumford and
Poldrack.

Rest: importance of confounds. Filtering + and detrending. Importance of
preprocessing.
Compcorr \cite{behzadi2007}
\cite{chai2011}
power, van dijk
\cite{satterthwaite2012}

FMRI signals contributing to functional connectivity
have been found to lie in frequencies below 0.1 Hz \cite{cordes2000}
XXX: but cite recent work that shows that there is info at higher
frequency.  Regressing out task??

@@

\begin{figure}
\newcommand{\photoframe}[1]{\setlength\fboxsep{0pt}%
{\color{black!70}\fbox{#1}\color{black}}}%
\begin{minipage}{.66\linewidth}%
    \includegraphics[width=\linewidth]{group_emp_cov.pdf}%
    \raisebox{.606\linewidth}{%
	\hspace*{-.462\linewidth}%
	\photoframe{%
	%\setlength\fboxsep{1pt}%
	\colorbox{white}{%
	\raisebox{.003\linewidth}{%
	    \includegraphics[width=.455\linewidth]{group_emp_cov_3d_crop}}%
	\rule{0pt}{.387\linewidth}%
	}}%
    }
\end{minipage}%
\hspace*{1pt}%
\begin{minipage}{.32\linewidth}%
    \includegraphics[width=\linewidth]{group_emp_cov_no_confounds.pdf}%

\caption{
Correlation matrices
Note that no filter was applied. Filtering low frequencies diminishes the
impact of some confounds that are removed ...
}
\end{figure}

\begin{figure}
\newcommand{\photoframe}[1]{\setlength\fboxsep{0pt}%
{\color{black!70}\fbox{#1}\color{black}}}%
\begin{minipage}{.66\linewidth}%
    \includegraphics[width=\linewidth]{group_l21_prec.pdf}%
    \raisebox{.606\linewidth}{%
	\hspace*{-.462\linewidth}%
	\photoframe{%
	%\setlength\fboxsep{1pt}%
	\colorbox{white}{%
	\raisebox{.003\linewidth}{%
	    \includegraphics[width=.455\linewidth]{group_l21_prec_3d_crop}}%
	\rule{0pt}{.387\linewidth}%
	}}%
    }
\end{minipage}%
\hspace*{1pt}%
\begin{minipage}{.32\linewidth}%
    \includegraphics[width=\linewidth]{group_emp_prec.pdf}%

@@

\caption{
Inverse-covariance matrices
}
\end{figure}

>>>>

BOLD signal during rest is typically just 5-10\% lower than during task-based
experiments (cite Raichle,  Mintun, Annu. Rev. Neurosci. 29, 449–476 (2006).)

@@

Define on-going vs evoked activity

@@

% XXX: discuss tradeoffs in the number of regions

@@

%{{{------ Parcellation figure -----------------------------------------------
\begin{figure}
\includegraphics[width=.5\linewidth]{aal_atlas.pdf}%
\hfill%
\includegraphics[width=.5\linewidth]{ho_atlas.pdf}%

\includegraphics[width=.5\linewidth]{sulci_atlas.pdf}%
\hfill%
\includegraphics[width=.5\linewidth]{ncuts_atlas.pdf}%

@@

\caption{
Different full-brain parcellations: the AAL atlas
\cite{tzourio-mazoyer2002a}, the Harvard-Oxford atlas, the sucli atlas used in
\cite{varoquaux2010c}, regions extracted by Ncuts
\cite{craddock2011}, the resting-state networks extracted in
\cite{smith2009} by ICA, and in \cite{varoquaux2011} by sparse dictionary
learning.
\label{fig:parcellations}
}
\end{figure}
%---------------------------------------------------------------------------}}}

@@

The concept of functional connectivity has be called elusive
\cite{horwitz2003}: it has many mathematical instantiations although in
essence they all strive to extract simple statistics from functional
imaging to characterize synchrony and communication between large
ensembles of neurons. Here we choose to focus on second order statistics
that can be related to Gaussian models, the simplest of which being the
correlation matrix of the signals of the different ROIs.

%{{{------ Correlation matrices figure ---------------------------------------
\begin{figure}
\newcommand{\photoframe}[1]{\setlength\fboxsep{0pt}%
{\color{black!70}\fbox{#1}\color{black}}}%
\begin{minipage}{.66\linewidth}%
    \includegraphics[width=\linewidth]{group_emp_cov.pdf}%
    \raisebox{.606\linewidth}{%
	\hspace*{-.462\linewidth}%
	\photoframe{%
	%\setlength\fboxsep{1pt}%
	\colorbox{white}{%
	\raisebox{.003\linewidth}{%
	    \includegraphics[width=.455\linewidth]{group_emp_cov_3d_crop}}%
	\rule{0pt}{.387\linewidth}%
	}}%
    }
\end{minipage}%
\hspace*{1pt}%
\begin{minipage}{.32\linewidth}%
    \includegraphics[width=\linewidth]{group_emp_cov_no_confounds.pdf}%

% XXX: need to give details on the dataset used here

\caption{
Correlation matrices of rest time-series extracted from the 39 main
regions of the Varoquaux 2011 \cite{varoquaux2011} parcellation -- 
Upper left: regressing out CompCorr signals, as well as white matter and
CSF average signals and movement parameters. The insert shows the
connections restricted to the default mode network.
% XXX: should I outline the default mode network on the main matrix?
-- Upper right: regressing out only movement parameters. -- Lower right:
regressing out movement parameters and global signal mean.
% XXX: should this remark go here?
No frequency filtering was applied here.
\label{fig:correlation_matrices}
}
\end{figure}
%---------------------------------------------------------------------------}}}

% The following paragraph is convoluted: we start to say that rest is
% from time-series, and task from beta, we go on discussing about beta
% maps, and then we move back to rest. It might need restructuring

% Horrible sentence below

\paragraph{Signal extraction}
%
Given ROIs, the next step is to extract from them a summary measure. To
study \emph{on-going} activity, \emph{e.g.} with rest data, signal
extraction is performed from the EPI time series; however, to study
\emph{evoked} activity with task-driven studies it can be beneficial to
run a first-level analysis, enforcing specificity of the measure
extracted to the task. With slow event-related design, task-specific
functional connectivity can be captured in trial-to-trial fluctuations in
the BOLD response, estimated using a GLM analysis with one regressor per
trial \cite{grillon2012,rissman2004}. This approach, known as beta-series
regression, has been adapted for rapid event-related designs, using
multiple GLMs to optimize deconvolution of each trial \cite{mumford2012}.
With rest datasets, it is important to regress out time series capturing
movement and other sources of structured noise, to separate on-going
activity from confounding signals. In addition to removing movement
parameters estimated during preprocessing, removing linear trends, signal
extracted from the white matter and the CSF \cite{chang2009}, as well as
using the CompCor \cite{behzadi2007} procedure gives more contrasted
correlation matrices that outline functional structures better
(fig.\,\ref{fig:correlation_matrices}). Filtering high frequencies is
often recommended, based on the initial observation that neural-related
signal are observed below 0.1\,Hz \cite{cordes2001,biswal1995}. While
high-pass and low-pass filtering decrease the impact of some confounds,
recent studies have shown that information on neural processes is present
in the full spectrum of frequencies observed
\cite{smith2012,vanoort2012}. Regressing out a good choice of confound
signals is more specific than frequency filter, and in our experience
gives more contrasted correlation matrices\footnote{Note that naive use
of filtering can induce spurious correlations \cite{davey2012}.}.

Finally, it is important to keep in mind that using resting-state to
study on-going activity is fragile to confounds such as movement
\cite{vandijk2012,power2011} or scanner noise, probably more so than
activation studies as it is harder to control the specificity of the
signal extracted. Special care must be taken in preprocessing strategies
\cite{vandijk2010,satterthwaite2012}.
%
In addition, improved specificity to BOLD signal can be enforced by
summarizing the signal in an ROI by a mean of the different voxels
weighted by the subject-specific grey matter probabilistic segmentation,
as output by \emph{e.g.} SPM's segmentation tool \cite{ashburner2005}, .

% Regressing out task?? Skipping for lack of space

@@

%{{{------ Precision matrices figure -----------------------------------------
\begin{figure}
\newcommand{\photoframe}[1]{\setlength\fboxsep{0pt}%
{\color{black!70}\fbox{#1}\color{black}}}%
\begin{minipage}{.66\linewidth}%
    \includegraphics[width=\linewidth]{group_l21_prec.pdf}%
    \raisebox{.606\linewidth}{%
	\hspace*{-.462\linewidth}%
	\photoframe{%
	%\setlength\fboxsep{1pt}%
	\colorbox{white}{%
	\raisebox{.003\linewidth}{%
	    \includegraphics[width=.455\linewidth]{group_l21_prec_3d_crop}}%
	\rule{0pt}{.387\linewidth}%
	}}%
    }
\end{minipage}%
\hspace*{1pt}%
\begin{minipage}{.32\linewidth}%
    \includegraphics[width=\linewidth]{group_emp_prec.pdf}%

@@

\caption{
Inverse-covariance matrices
}
\end{figure}
%---------------------------------------------------------------------------}}}

% XXX: should we have a summary paragraph commenting on the structure of
% the graphs and correlation matrices extracted.
% And also commenting on the meaning of regressions and sign of
% correlation coefficients (the whole global mean debate).
\cite{chai2011} (note: chai2011 used compcor)
